Compare Betahat vs Zscore Analysis

### 2014-10-15 paper/Rcode/CompareBetahatvsZscoreAnalysis.rmd
updated user interface for the way the EE and ES models are fitted

It is straightforward to use the ashr package to perform analysis under the ES model:
one simply replaces betahat with the standardized betahat, $\hat\beta_j$, and the standard errors
for these standarized betahat values with 1. It is also straightforward to compare
the two competing modelling assumptions (EE vs ES) by computing the log likelihood ratio, $log[p_{EE}(\hat\beta | s, \hat g_{EE})/p_{ES}(\hat\beta | s, \hat g_{ES})]$.

@@

Here is the EE analysis.
```{r}
  library(ashr)
  ash1.res = ash(betahat, betahat.se,method="fdr")
```

And here is how we can perform the ES analysis,
by replacing betahat with the standardized betahat, and replacing the standard error with 1:
```{r}
  ash2.res = ash(betahat/betahat.se, 1,method="fdr")
```
Note that the posterior distributions computed by ash will now be for the standardized
effects, and not for the effects $\beta_j$. However, the local false sign rate (lfsr) and
related quantities are unaffected by this because the standardized effect is positive if and only if the effect is positive.

Now we compare the EE vs ES models:
```{r}
  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se,df=NULL, zscore=FALSE) 
  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL, zscore=TRUE)
  loglikEE-loglikES
```

@@

To finish, we illustrate that if we generate data under the EE model then that model
is preferred by the log likelihood:
```{r}
  set.seed(1234)
  nsamp=1000
  betahat.se = rgamma(nsamp,1,1)
  beta = rnorm(nsamp) #simulate effects under the original assumption 
  betahat = rnorm(nsamp,beta,betahat.se)
  zscore = betahat/betahat.se
  pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
  ash1.res = ash(betahat, betahat.se,method="fdr")
  ash2.res = ash(betahat/betahat.se, 1,method="fdr")
  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se, df=NULL,zscore=FALSE) 
  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL,zscore=TRUE)
  loglikEE-loglikES
```
So here the log likelihood ratio is positive, indicating that the EE model is preferred.

>>>>

It is straightforward to use the ashr package to perform analysis under the ES model:
simply specify model="ES". (Internally, this causes ash to 
simply replaces betahat with the standardized betahat, $\hat\beta_j$, and the standard errors
for these standarized betahat values with 1.) It is also straightforward to compare
the two competing modelling assumptions (EE vs ES) by computing the log likelihood ratio, $log[p_{EE}(\hat\beta | s, \hat g_{EE})/p_{ES}(\hat\beta | s, \hat g_{ES})]$.

@@

Here is the EE analysis.
```{r}
  library(ashr)
  ash1.res = ash(betahat, betahat.se,method="fdr",model="EE")
```

And here is how we can perform the ES analysis,
by replacing betahat with the standardized betahat, and replacing the standard error with 1:
```{r}
  ash2.res = ash(betahat, betahat.se,method="fdr", model="ES")
```
Note that the posterior distributions computed by ash will now be for the standardized
effects, and not for the effects $\beta_j$. However, the local false sign rate (lfsr) and
related quantities are unaffected by this because the standardized effect is positive if and only if the effect is positive.

Now we compare the EE vs ES models:
```{r}
  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se,df=NULL, model="EE") 
  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL, model="ES")
  loglikEE-loglikES
```

@@

The above illustrates these ideas on simulations from the ES model. For comparison,
we now provide results for simulations under the EE model.
```{r}
  set.seed(1234)
  nsamp=1000
  betahat.se = rgamma(nsamp,1,1)
  beta = rnorm(nsamp) #simulate effects under the original assumption 
  betahat = rnorm(nsamp,beta,betahat.se)
  zscore = betahat/betahat.se
  pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
  ash1.res = ash(betahat, betahat.se,method="fdr",model="EE")
  ash2.res = ash(betahat, betahat.se, method="fdr",model="ES")
  loglikEE=ashr:::loglik.ash(ash1.res, betahat, betahat.se, df=NULL,model = "EE") 
  loglikES=ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL,model = "ES")
  loglikEE-loglikES
```
So here the log likelihood ratio is positive, indicating that the EE model is preferred
(which is expected since the data were generated under that model).

Finally, note that the likelihood should always be computed under the same
model as was used when applying ash. You'll get a warning if you do this wrong, although it will still run:
```{r}
  ashr:::loglik.ash(ash2.res,betahat, betahat.se, df=NULL,model = "EE")
```
